An Adolescent Mini-Neptune in the Kepler Field

### 2021-08-13 paper/ms.tex
update ms

% gaia target sample numbers
\newcommand{\nkinematic}{1{,}201} % deltalyr_kc19_cleansubset_withreddening.csv
% all from oneoff_calculate_rotation_numbers.py
\newcommand{\noriginal}{3{,}071} % Vizier; KC19 Theia 73 member count
\newcommand{\nwithtess}{924} % How many stars have TESS data?
\newcommand{\nkinwithtess}{391} % How many kinematically selected stars have TESS data?
\newcommand{\nkinwithtessandcrowding}{192} % How many kinematically selected stars have TESS data and meeting the crowding cutoff (nequal<=0, nclose<=1)?
\newcommand{\nkindefaultcleaning}{157} % How many of these meet the "defaultcleaning" criteria (LSP>0.1, Prot<15, and crowding cut)?
\newcommand{\nrotgood}{147} % nkindefaultcleaning, minus outliers.

@@

Of the \noriginal\ candidate \cn\ members reported by
\citet{kounkel_untangling_2019}, TESS data were available for
\nwithtess\ stars.  We extracted light curves using the nearest pixel
to each star, and regressed them against systematics using the causal
pixel model implemented in the \texttt{unpopular} package
\citep{hattorio_2021_cpm}.  We then measured candidate rotation
periods using a Lomb-Scargle periodogram
\citep{lomb_1976,scargle_studies_1982,astropy_2018}.  To enable cuts
on crowding, we queried the Gaia source catalog for stars within a
$21\farcs0$ radius of the target star (1 TESS pixel).  We recorded the
number of stars within the aperture with greater brightness than the
target star, and with brightness within 1.25 and 2.5 TESS magnitudes
of the target star.  {\bf We then supplemented these measurements with XXX Kepler
rotation periods (CITEP MCQUILLAN2014)}.

We cleaned the candidate TESS rotation period measurements through a
combination of automated and manual steps.  As a matter of scope, we
restricted our attention to the \nkinwithtess\ stars in
the spatial and kinematic proximity of Kepler\,1627, as discussed in
Appendix~\ref{app:kinematicselection}.  We also only focused on
stars for which no companions were known with a brightness exceeding one-tenth of
the target star within a $21\farcs0$ radius: there were 
\nkinwithtessandcrowding\ stars that met these requirements, and that
had TESS data available.
For plotting purposes we then imposed a selection based on the strength of the
rotation period signal itself: we required the Lomb Scargle power to
exceed 0.1, and the period to be below 15\,days.

The lower panel of Figure~\ref{fig:age} shows the resulting
\nkindefaultcleaning\ stars.  The majority of these stars fall below
the ``slow sequence'' of the Pleiades, consistent with a
gyrochronological age for the \cn\ below 100\,Myr.  Approximately 10 stars 
appear as outliers above the ``slow sequence''.  Assuming that they
are all false positives ({\it i.e.}, field interlopers), the rotation
period detection fraction would be
\nrotgood/\nkinwithtessandcrowding$\approx 77\%$.

@@

To summarize, the age of the \cn\ cluster based on the empirical
fitting procedure described in Section~\ref{sec:cmd} is
$35^{+6}_{-5}$\, Myr.  This procedure hinges on the age adopted
for IC\,2602, since it is the closest reference cluster in the color-magnitude diagram.
Gyrochronologically, while most of the G and K
dwarfs in the \cn\ have not yet converged to form a slow sequence,
their rotation distribution and that of the F dwarfs is consistent
with the same age.  The rotation periods of other
30\,Myr to 50\,Myr clusters ({\it e.g.}, IC\,2602, Collinder 135)
are indistinguishable
(S.~Douglas, priv{.}~comm{.}).  The most precise age for \sn\ and its
planet are therefore also $35^{+6}_{-5}$\, Myr, based on the star's spatial,
kinematic, and rotational connection to this larger ensemble, and the assumption that
\pn\ formed very quickly after the star itself.

% TODO:  update this based on HIRES instead of TRES
% TODO:  cite Stephanie instead of priv comm? Send her your Prot
Beyond the kinematic and rotational lines of evidence for the youth of \sn,
there is a final consistency check: the abundance of Li\,\textsc{I} as
inferred from the 6708\,\AA\ doublet.  This connection of this element
to age measurements for young stars has been reviewed by
\citet{soderblom_ages_2014}.  Following the equivalent width
measurement procedure described by \citet{bouma_2021_ngc2516}, we find
${\rm EW}_{\rm Li} = 232^{+23}_{-13}\,$m\AA, which does not
correct for the Fe\,\textsc{I} blend at 6707.44\AA\.  Given $T_{\rm eff} \approx
5500\,{\rm K}$, this measurement is also in agreement with an age for
the star between 30\,Myr and 50\,Myr (see {\it e.g.}, Figure~9 of
\citealt{bouma_2021_ngc2516}, or the data
from \citealt{randich_gaiaeso_2018}).

>>>>

% gaia target sample numbers
\newcommand{\nkinematic}{1{,}201} % deltalyr_kc19_cleansubset_withreddening.csv
% all from oneoff_calculate_rotation_numbers.py
\newcommand{\noriginal}{3{,}071} % Vizier; KC19 Theia 73 member count
\newcommand{\nwithtess}{924} % How many stars have TESS data?
\newcommand{\nkinwithtess}{391} % How many kinematically selected stars have TESS data?
\newcommand{\nkinwithtessandcrowding}{192} % How many kinematically selected stars have TESS data and meeting the crowding cutoff (nequal<=0, nclose<=1)?
\newcommand{\nkindefaultcleaning}{145} % How many of these meet the "defaultcleaning" criteria (LSP>0.1, Prot<15, and crowding cut)?
\newcommand{\nrotgood}{135} % nkindefaultcleaning, minus outliers.
\newcommand{\nfracprot}{70} % fraction nrotgood/nkinwithtessandcrowding

@@

Of the \noriginal\ candidate \cn\ members reported by
\citet{kounkel_untangling_2019}, TESS data were available for
\nwithtess\ stars.  \footnote{Kepler rotation periods were derived by
\citet{mcquillan_rotation_2014} for 56 stars; in these cases, we
adopted the Kepler rotation period.  } We extracted light curves from
the TESS images using the nearest pixel to each star, and regressed
them against systematics with the causal pixel model implemented in
the \texttt{unpopular} package \citep{hattorio_2021_cpm}.  We then
measured candidate rotation periods using a Lomb-Scargle periodogram
\citep{lomb_1976,scargle_studies_1982,astropy_2018}.  To enable cuts
on crowding, we queried the Gaia source catalog for stars within a
$21\farcs0$ radius of the target star (1 TESS pixel).  We recorded the
number of stars within the aperture with greater brightness than the
target star, and with brightness within 1.25 and 2.5 TESS magnitudes
of the target star.

We cleaned the candidate TESS rotation period measurements through a
combination of automated and manual steps.  As a matter of scope, we
restricted our attention to the \nkinwithtess\ stars in the spatial
and kinematic proximity of Kepler\,1627, as discussed in
Appendix~\ref{app:kinematicselection}.  We also only focused on stars
for which no companions were known with a brightness exceeding
one-tenth of the target star within a $21\farcs0$ radius: there were
\nkinwithtessandcrowding\ stars that met these requirements, and that
had TESS data available.  For plotting purposes we then imposed a
selection based on the strength of the rotation period signal itself:
we required the Lomb Scargle power to exceed 0.2, and the period to be
below 15\,days.

The lower panel of Figure~\ref{fig:age} shows the resulting
\nkindefaultcleaning\ stars.  The majority of these stars fall below
the ``slow sequence'' of the Pleiades, consistent with a
gyrochronological age for the \cn\ below 100\,Myr.  Approximately 10
stars appear as outliers above the ``slow sequence''.  Assuming that
they are all false positives ({\it i.e.}, field interlopers), the
rotation period detection fraction would be
\nrotgood/\nkinwithtessandcrowding$\approx \nfracprot\%$.

@@

To summarize, the age of the \cn\ cluster based on the empirical
fitting procedure described in Section~\ref{sec:cmd} is
$35^{+6}_{-5}$\,Myr.  This procedure hinges on the age adopted for
IC\,2602, since it is the closest reference cluster in the
color-magnitude diagram.  Gyrochronologically, while most of the G and
K dwarfs in the \cn\ have not yet converged to a slow sequence, their
rotation distribution and that of the F dwarfs is consistent with the
same age.  The rotation periods of other 30\,Myr to 50\,Myr clusters
({\it e.g.}, IC\,2602 and Collinder 135) are indistinguishable
(S.~Douglas, priv{.}~comm{.}).  The most precise age for \sn\ and its
planet are therefore also $35^{+6}_{-5}$\, Myr, based on the star's
spatial, kinematic, and rotational connection to this larger ensemble,
and the assumption that the planet formed very quickly after the star.

Beyond the kinematic and rotational lines of evidence for the youth of
\sn, there is a final consistency check: the abundance of
Li\,\textsc{I} as inferred from the 6708\,\AA\ doublet.  The utility
of this element for age dating young stars has been reviewed by
\citet{soderblom_ages_2014}.  We acquired an iodine-free spectrum from
Keck/HIRES on the night of 26 March 2021 using the standard setup and
reduction techniques of the California Planet Survey
\citep{howard_cps_2010}.  Following the equivalent width measurement
procedure described by \citet{bouma_2021_ngc2516}, we measure ${\rm
EW}_{\rm Li} = 233^{+5}_{-7}\,$m\AA.   Note that this value does not
correct for the Fe\,\textsc{I} blend at 6707.44\AA\.  Given $T_{\rm
eff} \approx 5500\,{\rm K}$, this measurement is also in agreement
with an age for the star between 30\,Myr and 50\,Myr (see {\it e.g.},
Figure~9 of \citealt{bouma_2021_ngc2516}, or the data from
\citealt{randich_gaiaeso_2018}).
